From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-23 Pi+/paper/finite.tex
Revise Lehmer subsection

\begin{proposition}
  For all $n : \Nat$, \( \Sn \eqv \im{\normf} \eqv \Lehmer[n] \).
\end{proposition}

@@

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item For any $k : \Fin[n]$, $\unit \sqcup \FinExcept{n}[k] \eqv \Fin[n]$.
    \item For any $k : \Fin[\suc[n]]$, $\FinExcept{\suc[n]}[k] \eqv \Fin[n]$.
    \item For any $n : \Nat$,
          \( \Aut[\Fin[\suc[n]]] \eqv \dsum{k : \Fin[\suc[n]]}{\FinExcept{\suc[n]}[\fzero] \eqv \FinExcept{\suc[n]}[k]} \).
  \end{enumerate}
\end{proposition}

@@

  The third proposition performs some interesting computations. On the left, we have the type of automorphisms of
  $\Fin[\suc[n]]$. Assume a particular one $\phi : \Fin[\suc[n]] \xrightarrow{\sim} \Fin[\suc[n]]$. To construct an
  element of the pair on the right, we pick the first component to be the image of the element $\zero$ under $\phi$.
  After fixing $k \defeq \phi(\zero)$, we need to compute the bijection in the second component. We remove the two
  elements $\zero$ and $k$ from $\Fin[\suc[n]]$ and construct an equivalence between the resulting types
  $\FinExcept{\suc[n]}[\zero]$ and $\FinExcept{\suc[n]}[k]$ as follows:
  $\FinExcept{\suc[n]}[\zero] \eqv \dsum{k:\Fin[\suc[n]]}{(\phi(\zero) \neq \phi(k))} \eqv \FinExcept{\suc[n]}[k]$

@@

To remind ourselves, in a Lehmer code, the element $\el{r}$ at postition $k$ counts the number of inversions of the
element $\el{r}$. This corresponds exactly to what happens in the third proposition -- inserting element $\el{r}$ at the
position $k$ shifts it by $k$ places.

\begin{proposition}
  For all $n:\Nat$, \( \Lehmer[n] \eqv \Aut[\Fin[\suc[n]]] \).
\end{proposition}

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]] \\
    \end{array}
    \qquad\qquad
    \begin{array}{rl}
           & \Lehmer[\suc[m]]                                                                              \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                         \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                              \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[\fzero] \eqv \Fin[\suc[m]]}          \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[\fzero] \eqv \FinExcept{\suc[m]}[k]} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]
    \end{array}
  \end{gather*}
\end{proof}

>>>>

\begin{proposition}~\label{prop:sn-im-lehmer-equiv}
  For all $n : \Nat$, \( \Sn \eqv \im{\normf} \eqv \Lehmer[n] \).
\end{proposition}

@@

\begin{proposition}~\label{prop:fin-finexcept}
  \leavevmode
  \begin{enumerate}
    \item For any $k : \Fin[n]$, $\unit \sqcup \FinExcept{n}[k] \eqv \Fin[n]$.~\label{prop:fin-finexcept-1}
    \item For any $k : \Fin[\suc[n]]$, $\FinExcept{\suc[n]}[k] \eqv \Fin[n]$.~\label{prop:fin-finexcept-2}
    \item For any $n : \Nat$,
          \( \Aut[\Fin[\suc[n]]] \eqv \dsum{k : \Fin[\suc[n]]}{\FinExcept{\suc[n]}[\fzero] \eqv \FinExcept{\suc[n]}[k]} \).~\label{prop:fin-finexcept-3}
  \end{enumerate}
\end{proposition}

@@

  The third proposition performs some interesting computations. On the left, we have the type of automorphisms of
  $\Fin[\suc[n]]$. Assume a particular one $\phi : \Fin[\suc[n]] \xrightarrow{\sim} \Fin[\suc[n]]$. To construct an
  element of the pair on the right, we pick the first component to be the image of the element $\zero$ under $\phi$.
  After fixing $k \defeq \phi(\zero)$, we need to compute the bijection in the second component. We remove the two
  elements $\zero$ and $k$ from $\Fin[\suc[n]]$ and construct an equivalence between the resulting types
  $\FinExcept{\suc[n]}[\zero]$ and $\FinExcept{\suc[n]}[k]$ as follows:
  $\FinExcept{\suc[n]}[\zero] \eqv \dsum{k:\Fin[\suc[n]]}{(\phi(\zero) \neq \phi(k))} \eqv \FinExcept{\suc[n]}[k]$.

@@

Using these facts, we can now prove the main result of this section.

\begin{proposition}~\label{prop:lehmer-aut-equiv}
  For all $n:\Nat$, \( \Lehmer[n] \eqv \Aut[\Fin[\suc[n]]] \).
\end{proposition}

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]] \\
    \end{array}
    \qquad\qquad
    \begin{array}{rlr}
           & \Lehmer[\suc[m]]                                                                              &                                                                 \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                         & \text{by definition}                                            \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                & \text{induction hypothesis}                                     \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                              & \text{$\Sigma$ over a constant family}                          \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[\fzero] \eqv \Fin[\suc[m]]}          & \text{by~\cref{prop:fin-finexcept}.~\ref{prop:fin-finexcept-2}} \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[\fzero] \eqv \FinExcept{\suc[m]}[k]} & \text{by~\cref{prop:fin-finexcept}.~\ref{prop:fin-finexcept-2}} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]                                                                     & \text{by~\cref{prop:fin-finexcept}.~\ref{prop:fin-finexcept-3}} \\
    \end{array}
  \end{gather*}
\end{proof}

\jk{TODO an example here}
To remind ourselves, in a Lehmer code, the element $\el{r}$ at postition $k$ counts the number of inversions of the
element $\el{r}$. This corresponds exactly to what happens in the \cref{prop:fin-finexcept-3}. Looking at the operation
combinatorially, it is inserting element $\el{r}$ at the position $k$, meaning that it shifts the element
$k$places.\jk{needs rewriting after the example is here}

By composing~\cref{prop:lehmer-aut-equiv} and~\cref{prop:sn-im-lehmer-equiv}, we obtain the final
equivalence.
